---
title: "GAPMINDER|LIFE EXPECTANCY 2007"
format: dashboard
theme: Lux
---

```{r}
pacman::p_load('tidyverse', 'gapminder', 'bslib','bsicons',
               'shiny', 'rnaturalearth', 'countrycode', 'plotly',
               'htmltools', 'reactable')
```

```{r}
gap_07 <- gapminder %>% 
  filter(year==2007)
highest_lifeExp_country <- gap_07 %>% 
  arrange(-lifeExp) %>% 
  head(1) %>% 
  pull(country)
highest_lifeExp_value <- gap_07 %>% 
  arrange(-lifeExp) %>% 
  head(1) %>% 
  pull(lifeExp) %>% 
  round(1)
lowest_lifeExp_country <- gap_07 %>% 
  arrange(lifeExp) %>% 
  head(1) %>% 
  pull(country)
lowest_lifeExp_value <- gap_07%>% 
  arrange(lifeExp) %>% 
  head(1) %>% 
  pull(lifeExp) %>% 
  round(1)
avaerage_lifeExp <- gap_07 %>% 
  summarise(weighted_mean=sum(lifeExp*pop)/sum(pop)) %>% 
  round(1)
  
```


# HOME


## Row 1 {height=25%}

```{r}
value_box(
  title = "Highest life expectancy",
  value = "82.6",
  showcase = bsicons::bs_icon("thermometer-high"),
  theme = value_box_theme(bg="#508beb"),
  p(paste0("(", highest_lifeExp_country, ")")),
  p("Japan")
)
```


```{r}
value_box(
  title = "Lowest life expectancy",
  value = "39.6",
  showcase = bsicons::bs_icon("thermometer-low"),
  theme = value_box_theme(bg="#274370"),
  p(paste0("(", lowest_lifeExp_country, ")"))
)
```

```{r}
value_box(
  title = "Average life expectancy",
  value = "68.9",
  showcase = bsicons::bs_icon("symmetry-horizontal"),
  theme = value_box_theme(bg="#919eb3")
)
```


## Row 2 {height=75%}


### {width=70%}

```{r}
#| include: false
country_shapes <- rnaturalearth::ne_countries()
ggplot(country_shapes) +
  geom_sf()

country_shapes
gap_07 <- gap_07 %>% 
  mutate(country_codes = countrycode(country,
                                     "country.name", "iso3c"))

gap_map <- left_join(country_shapes, gap_07,
          by=c("adm0_a3"="country_codes")) %>% 
  mutate(tooltip_label = paste(country, round(lifeExp, 1),
                               sep = ": ")) %>% 
  ggplot() +
  geom_sf(aes(fill = lifeExp, text = tooltip_label)) +
  theme_void() + 
  theme(legend.position = 'none')
  
gap_map_ggplotly <- ggplotly(gap_map, tooltip = "text")

```


```{r title = "Countries Map with Life Expectancy"}
gap_map_ggplotly
```


### {width=30%}

```{r}
top_20_countries <- gap_07 %>% 
  arrange(desc(lifeExp)) %>% 
  head(20) %>% 
  mutate(tooltip_label = paste(country, round(lifeExp, 1),
                               sep = ": ")) %>% 
  ggplot(aes(y= reorder (country, lifeExp), x = lifeExp,
             fill = lifeExp,
             text = tooltip_label)) +
  geom_col() +
  geom_text(aes(label = round(lifeExp, 1)),
            nudge_x = -10,
            color = "white") +
  labs(y = "country", x = "Life Exp") +
  theme(legend.position = "none")

top_20_countries_ggplotly <-
  ggplotly(top_20_countries, tooltip = 'text')

```


```{r title="Top 20 by LifeExp"}
top_20_countries_ggplotly
```



# DOWNLOAD DATA

The data used in this dashboard is shown below and can be downloaded
as a CSV.

```{r}
library(htmltools)
htmltools::browsable(
  tagList(
    reactable(gapminder,
              elementId = "gapminder-table",
              searchable = T,
              filterable = T),
    
tags$button("Download as CSV", onclick=
                        "Reactable.downloadDataCSV('gapminder')")))
```


# ABOUT

This data comes from the r 'gapminder' package, and originally sourced
from the Gapminder Foundation.

Gapminder Foundation is a non-profit venture registered in Stockholm, Sweden, that promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic, and environmental development at local, national, and global levels.

An example of one of videos of Hans Rosling's is show here below:

<iframe width="560" height="315" src="https://www.youtube.com/embed/hVimVzgtD6w?si=JXczr1NDqAqnoVEx" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>





